%  ========================================================================
%  Copyright (c) 1985 The University of Washington
%
%  Licensed under the Apache License, Version 2.0 (the "License");
%  you may not use this file except in compliance with the License.
%  You may obtain a copy of the License at
%
%      http://www.apache.org/licenses/LICENSE-2.0
%
%  Unless required by applicable law or agreed to in writing, software
%  distributed under the License is distributed on an "AS IS" BASIS,
%  WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
%  See the License for the specific language governing permissions and
%  limitations under the License.
%  ========================================================================
%    Printed in twoside style now that that's allowed
%
 
\documentclass[12pt, twoside]{uwthesis}
 
% The following line would print the thesis in a postscript font 
\usepackage{xspace}

%\usepackage[sectionbib]{natbib}

\usepackage[sort&compress]{natbib}  % square,numbers,
\usepackage{chapterbib}
\usepackage{multibib}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{float}
\usepackage{xspace}
\usepackage{hyperref}
\usepackage{aas_macros}
\usepackage{appendix}
\usepackage{etoolbox}
\usepackage{longtable}
\usepackage{listings}
\usepackage{color}
\usepackage{inconsolata}
\usepackage{sidecap}

\usepackage[footnote,printonlyused]{acronym}
\usepackage{scrextend}
\deffootnote[1.0em]{1em}{0em}{\textsuperscript{\thefootnotemark}\,\enskip}

\newcommand{\kepler}{\textsl{Kepler}\xspace}
\newcommand{\astroplan}{\texttt{astroplan}\xspace}
\newcommand{\astropy}{\texttt{astropy}\xspace}
\newcommand{\pyephem}{\texttt{pyephem}\xspace}
\newcommand{\stsp}{\texttt{STSP}\xspace}
\newcommand{\gaia}{\textit{Gaia}\xspace}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\renewcommand{\lstlistingname}{Example}
\lstset{frame=tb,
  language=Python,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3,
}

%\newcommand\ion[2]{#1$\;${%
%\ifx\@currsize\normalsize\small \else
%\ifx\@currsize\small\footnotesize \else
%\ifx\@currsize\footnotesize\scriptsize \else
%\ifx\@currsize\scriptsize\tiny \else
%\ifx\@currsize\large\normalsize \else
%\ifx\@currsize\Large\large
%\fi\fi\fi\fi\fi\fi
%\rmfamily\@Roman{#2}}\relax}% 

%\def\bibpreamble{\protect\addcontentsline{toc}{chapter}{Bibliography}}

%\AtBeginEnvironment{subappendices}{%
%  \addtocontents{toc}{\protect\addvspace{10pt}Appendices}
%}

\makeatletter
\patchcmd{\@chapter}{\protect\numberline{\thechapter}#1}
{\@chapapp~\thechapter: #1}{}{}
\makeatother

\setcounter{tocdepth}{1}  % Print the chapter and sections to the toc


\begin{document}
 
% ==========   Preliminary pages
%
% ( revised 2012 for electronic submission )
%

\prelimpages
%
% ----- copyright and title pages
%
\Title{Photometric, Spectroscopic and Astrometric Perspectives on Stellar Magnetic Activity, and Its Effect on Exoplanet Characterization}
\Author{Brett M. Morris}
\Year{2019}
\Program{Astronomy and Astrobiology}

\Chair{Eric Agol}{}{Astronomy Department}
\Signature{Suzanne L. Hawley}

\copyrightpage

\titlepage


\setcounter{page}{-1}
\abstract{%
Exoplanets have been discovered principally via the transit method, which reveals planetary radii and orbital periods. In transiting multi-planet systems, we can measure the masses of exoplanets without expensive spectroscopy by using transit timing variations. The orbit of a single transiting planet around a single star would be perfectly periodic, but if there's more than one planet in the system, the gravitational influence of each planet on each other pulls the planets slightly ahead or behind in their orbits. The apparent early or late arrival of an exoplanet transit thus transmits information about the mass of the perturbing planet. Transits can also reveal the composition of exoplanet atmospheres via transmission spectroscopy. When the planet passes in front of the host star, the planet will appear largest at wavelengths where the planet's atmosphere is opaque, and smallest at wavelengths where the atmosphere is transparent. Thus by measuring the apparent radius of the planet as a function of wavelength, we can attain a spectrum of an exoplanet's atmosphere. However, there are confounding signals generated by magnetic activity near the stellar surface which inject confounding time- and wavelength-dependent signals into the spectrophotometry of exoplanet host stars which complicate all of the aforementioned measurements. 

The goal of this thesis is to explore stellar magnetic activity and its impact on measurements of exoplanet properties. I used the transiting planet HAT-P-11 b to measure the sizes and latitude distribution of starspots on its active K4 dwarf host star, and find that its magnetic activity mirrors the Sun's. I measured the chromospheric activity of HAT-P-11 and compared it to stars of similar temperature and rotation period and find that it's the most active of its peers, perhaps suggestive of star-planet interaction. I directly measured starspot coverage on a sample of bright stars via TiO molecular band modeling. With ground-based photometry, I measured the transit times of TRAPPIST-1 b and c, and several Kepler host stars. I devised a technique for measuring robust exoplanet radii even in the presence of significant starspot distributions by reparameterizing the transit light curve of Mandel \& Agol (2002). I identified the possibility of detecting stellar activity cycles of nearby stars using precision astrometry. Finally, I devised a simulator for James Webb Space Telescope observations of transiting exoplanets, to explore the limits imposed by stellar magnetic activity on transit timing and transmission spectroscopy measurements. 
}
 
%
% ----- contents & etc.
%
\tableofcontents
%\listoffigures
%\listoftables  % I have no tables
 

% ----- glossary 
%
% \chapter*{Glossary}      % starred form omits the `chapter x'
% \addcontentsline{toc}{chapter}{Glossary}
% \thispagestyle{plain}
% %
% \begin{glossary}
% \item[argument] replacement text which customizes a \LaTeX\ macro for
% \end{glossary}
 
% ----- acknowledgments
\acknowledgments{% \vskip2pc
There are not enough pages in the world to properly acknowledge the people who made this dissertation possible. The author first needs to recognize the unrelenting, unconditional support of Shari, Scott and Michael Morris. 

The author had the privilege of being taught science by the following teachers at Half Hollow Hills Central School District in New York: Mary Ann and Richard Sloan, Elaine Tozar, Janet Chan, Paul Travaglianti, Rick Russo, Thomas Page, Chrysi Notskas, Rose Suarez, and Dr.~Michael Lake. The author owes special thanks to Thomas Carey and Thomas Affatigato, who gave the author a cathedral to work in; and to Dennis Grunbeck, who introduced the author to physics.

At the University of Maryland, the author was fortunate enough to have the guidance of Professors Chris Reynolds, Derek Richardson, who demonstrated infinite patience and incredible encouragement, and L.~Drake Deming, who skillfully and joyfully transmitted invaluable skills, and gives special thanks to Elizabeth Warner for trusting him with her telescope. 

At the University of Washington, Professors Eric Agol and Suzanne L. Hawley stewarded the author through challenges in life as well as science. Their patience, understanding and encouragement made this work possible.  
}
%
% ----- dedication
%
\dedication{\begin{center}
\vspace{1 cm}To Rebecca, my best friend, who came with me on this adventure; 

\vspace{1cm} to my family, who sent me out on it;

\vspace{1cm} and to my friends, who were patient while I was gone.
\end{center}}

%
% end of the preliminary pages

%
% ==========      Text pages
%

\textpages

\part{Solar Variability and Magnetic Activity} \label{part:solaractivity}

\begin{quote}
{\it Starting from fish-shape Paumanok where I was born...\\
How curious! how real!\\
Underfoot the divine soil, overhead the sun.}
\hfill --- Walt Whitman
\end{quote}

For the foreseeable future, an astronomer's only available tool for studying an exoplanet, or a planet outside of the solar system, is light. That light may be light that disappears from the host star during an exoplanet transit, transmitting information via the planet's silhouette (explanations of each of these techniques will follow in Chapter~\ref{chap:intro}). It can be starlight that has filtered through the planet's atmosphere as is used in transmission spectroscopy. It might be starlight that has reflected off of the planet in the case of the secondary eclipse or direct imaging. 

Each of these techniques make use of the light of the exoplanet's host star to study the properties of the exoplanet. Our reliance on stars to illuminate their exoplanets has led to the oft-repeated mantras ``one only knows as much about a planet as one knows about its host star,'' and ``know thy star, know thy planet'' (attributed to David Ciardi). If we are to have any hope of understanding the detailed properties of exoplanets, we must first thoroughly understand the light sources which make them visible to us. 

Stars are not pristine light sources -- they change in brightness with time, they have surface brightness variations like dark starspots or bright facular regions, and each of these surface components have different spectral properties. The deviations from pristine light sources are often generated by strong magnetic fields affecting the stellar plasma, and therefore are generally grouped into the broad category of ``stellar magnetic activity''. 

In this Part, we introduce the subject matter of this thesis in Chapter~\ref{chap:intro}, with particular attention paid to the activity of our nearest stellar neighbor, the Sun. 

\chapter{Introduction}  \label{chap:intro}

\input{intro/introduction.tex}

\section{Brief Introduction to the Magnetohydrodynamics of Starspots}

Sunspots have been observed on the Sun with the naked eye for millennia \citep{Hayakawa2017}, and with the telescope since Galileo \citep{Galilei1613}. The observation that spots are highly magnetized was first made by \citet{Hale1908}, which lead him and colleagues, notably Sir John Larmor, to question ``how could a rotating body such as the Sun become a magnet?'' \citep{Larmor1919}. This question is a vexing one as it requires three-dimensional magnetohydrodynamics to solve. In the following section, we will briefly outline the mechanism by which starspots arise. 

The plasma in the outer convective zones of main sequence stars is conducting and magnetized, so its dynamics are dictated by Maxwell's equations: 
\begin{eqnarray}
\nabla \times \vec{E} &=& -\frac{\partial \vec{B}}{\partial t}\\
\nabla \times \vec{B} &=& \mu_0 \vec{J}\\
\vec{J}/\sigma &=& \vec{E} + \vec{v} \times \vec{B}\\
\frac{\partial \vec{B}}{\partial t} &=& \nabla \times \left( \vec{v} \times \vec{B} - \eta \nabla \times \vec{B} \right)
\end{eqnarray}
Where $\vec{E}$ and $\vec{B}$ are the electric and magnetic field vectors, $\vec{J}$ is the current vector, $\vec{v}$ is the plasma velocity, and $\eta$ is the magnetic diffusivity. The last relation, known as the induction equation, determines how the magnetic field evolves in time. On the right hand side, the first term, $\vec{v} \times \vec{B}$ is a source of inductance from bulk flow of the plasma, while the second term $\eta \nabla \times \vec{B}$ is a dissipative loss term due to resistance in the plasma. Using order of magnitude characteristic velocities (km/s convective motion), convective length scales (the solar radius) and resistivity, we find that the magnetic Reynolds number Rm:
\begin{eqnarray}
\textrm{Rm} = \frac{\textrm{induction}}{\textrm{dissipation}} = \frac{v L}{\eta}\\
v \sim 10^3\, \textrm{m s}^{-1}\\
L \sim R_\odot = 7 \times 10^8 \textrm{m}\\
\eta \sim 10^{0-5} \textrm{m$^2$ s}\\
\textrm{Rm} \sim 10^{6 - 10} \gg 1
\end{eqnarray}
\citep{Ossendrijver2003}. Thus in stars, the very high conductivity $\eta \rightarrow 0$ makes the dissipation term in the induction equation insignificant. In this regime we can apply the Alfv{\'e}n's ``frozen flux theorem'' \citep{Alfven1942}, which indicates that bulk flow of the plasma moves magnetic fields lines with it. This plays a critical role in the dynamics of plasma at the base of the convective zones in stars. The frozen flux theorem states that flux is conserved within an area $A$ before ($i$) and after ($f$) some stretching of the magnetic flux tube  
\begin{equation}
(B dA)_i = (B dA)_f,
\end{equation}
which can be combined with the conservation of mass for plasma of density $\rho$ in a flux tube of length $\ell$
\begin{equation}
(\rho \ell dA)_i = (\rho \ell dA)_f,
\end{equation}
to show that $B \propto \rho \ell$, or in words the magnetic field increases linearly with increasing length of the flux tube.

Imagine a tube of plasma with magnetic field lines running along the length of it, at the base of the convective zone. Sun-like stars are differentially rotating -- the Sun, for instance, rotates once in 25 days at the equator, and closer to 34 days at the poles. Differential rotation of the star will stretch the plasma tube, making it longer and narrower, and since the frozen flux theorem applies, the plasma will drag the magnetic field lines with it, bringing the field lines closer together, i.e.\ amplifying the magnetic field in the flux tube. The pressure equilibrium between the plasma in the interior of the flux tube and the exterior medium is given by the plasma pressure in each region plus the magnetic pressure, $B^2/2\mu_0$, and using the ideal gas law, we can solve for the difference in density between the magnetic flux tube and its surroundings assuming $B_{\rm int} \gg B_{\rm ext}$:
\begin{eqnarray}
p_{\rm ext} + \frac{B_{\rm ext}^2}{2\mu_0} &=& p_{\rm int} + \frac{B_{\rm int}^2}{2\mu_0}\\
P &=& \rho \frac{kT}{m} \\
\rho_{\rm ext} - \rho_{\rm int} &=& \frac{B_{\rm int}^2}{2\mu_0} \frac{m}{kT}
\end{eqnarray}
We note that the term on the right hand side of the last equation is always positive for $B_{\rm int} > 0$, and thus the magnetized plasma will have a lower density than its surroundings, causing it to become buoyant and to rise. 

In this classical picture, a portion of the rising magnetic flux tube now rises through the convective zone, where it is subject to stochastic convective turbulence and the consistent twist of the Coriolis force, before it finally rises through the photosphere \citep{Rempel2011, Cheung2014}. The magnetic field is very strong at the two locations where the flux tube intersects the photosphere, giving rise to two regions where the local magnetic field dominates over convection. These are called bipolar magnetic regions (BMRs), as observations of the magnetic polarity of each region show opposite polarity, known as Hale's law \citep{Hale1925} -- a natural result of a single flux tube exiting (outward field) and entering (inward field) the photosphere. The halves of the BMRs are referred to as the leading and following components according to the direction of rotation. The twist of the Coriolis force causes BMRs to have a slight tilt with respect to a line of latitude, with the following component at higher latitudes than the leading component in the northern hemisphere (known as Joy's Law). Since the upwelling of hot plasma from convection is suppressed by the strong magnetic fields, an area just below the quiescent photosphere is evacuated, pushing the effective photosphere downwards, known as the Wilson depression \citep{Wilson1774}. In regions of strong enough fields, the evacuated region cools and thus the cores of the BMRs are observed as dark sunspots. 

In summary, dark starspots are the magnetic footprints of strong magnetic fields, which originate at the base of the convective zone and rise through the photosphere due to magnetic buoyancy in differentially rotating stars. Spots often occur in pairs of opposite polarity known as bipolar magnetic regions which are slightly tilted with respect to lines of latitude.

\section{Brief Introduction to the Solar Dynamo} \label{sec:solardynamo}

In the previous section, we described the process by which a plasma tube with a magnetic field may rise through the surface and produce starspots observed in the photosphere. There is a problem lurking in our description of stellar magnetic fields -- stellar plasma is not in fact dissipationless, and in time, one might expect the bulk flows that give rise to amplified magnetic fields to decay due to resistance until fields are no longer produced. Or another challenge is the subject of how to construct a bulk flow of the plasma which remains magnetized, as it must not be axisymmetric in order to be sustained over millennia (known as Cowling's Theorem, see \citealt{Cowling1933}). In this section we'll briefly outline the mechanism known as the $\alpha  \Omega$ dynamo, which reproduces the observed solar magnetic activity cycle \citep{Hathaway2010, Charbonneau2014}. 

Suppose we start the Sun with a strictly dipolar field, like a bar magnet. The real Sun does have a dipolar field at solar minimum. Field lines in the convective zone run along meridional lines in this initial condition -- or in other words the field is strictly poloidal. Since the Sun is differentially rotating, and the frozen flux theorem applies, the field lines near the equator are dragged ahead of the field lines near the poles. After several rotations, the originally meridional field lines have been twisted around the Sun several times, wound up like a spring at the mid-latitudes, until a significant component of the magnetic field is running along lines of latitude -- or in other words, there is a significant toroidal component of the magnetic field. The twisting of the meridional (poloidal) magnetic field lines into azimuthal (toroidal) ones due to differential rotation is known as the $\Omega$-effect. 

Now we have the magnetic field lines wrapped around the star and concentrated (amplified) at the mid-latitudes. As differential rotation continues to stretch the plasma and amplify the magnetic field, the flux tubes eventually become unstable and rise through the photosphere, producing starspots via the mechanism described in the previous section. Some of the magnetic flux tubes have a twist imparted on them by the Coriolis effect and by convective turbulence \citep{Parker1955a, Parker1955b}, which together create loops in the twisted magnetic flux tubes. Magnetic reconnection occurs, liberating the magnetic field loops, which due to the earlier twist now contain a poloidal component. This is known as the $\alpha$-effect. 

Now the simultaneously present toroidal and poloidal field components interact with each other, causing bands of magnetic flux tubes to migrate from the mid-latitudes towards the equator, creating the observed dynamo wave of starspots that emerge first at mid-latitudes (near 30$^\circ$) and later near the equator throughout the solar activity cycle \citep{Parker1955a, Parker1955b}. It has been observed that meridional flow exists on the solar surface which flows from the equator to the poles at the surface, and via the continuity it is thought that meridional flow must also exist at the base of the convective zone from the poles to the equator. The inferred flow at the base of the convection zone is slow, about 1 m/s, which is roughly the speed needed for bulk flow of plasma to migrate from 30$^\circ$ latitude to the equator in 11 years -- causing some to speculate whether flux transport via the meridional flow is responsible for the 11 year periodicity of the solar activity cycle \citep{Dikpati2006,Hathaway2010}. 

\chapter{The Solar Benchmark: Rotational Modulation of the Sun Reconstructed from Archival Sunspot Records}

\input{howard/main.tex}

\chapter{The Stellar Oscillation Noise Floor for Transiting Exoplanet Photometry with PLATO}

\chapter{Generating Stellar Light Curves with Granulation and Oscillations}


\part{Stellar Magnetic Activity} \label{part:activity}

\begin{verse}
{\it I believe a leaf of grass is no less than the journey work of the stars.\\}
\hfill --- Walt Whitman
\end{verse}

In Chapters~\ref{chap:h11} and \ref{chap:h11sindex} we present a detailed investigation into the properties of stellar activity of the star HAT-P-11, a star with a similar rotation period to the Sun, but 20\% less massive. In Chapter~\ref{chap:freckles} we experiment with a technique for measuring the presence and extent of stellar magnetic activity via stellar spectroscopy, and in Chapter~\ref{chap:gaia} we introduce a novel technique for measuring stellar activity using astrometry.  In Chapter~\ref{chap:nephelion} we investigate whether a handful of Sun-like stars exhibit solar-light magnetic activity. Finally, in Chapter~\ref{chap:trappist1_bright} we probe the magnetic activity of a star very different from the sun, TRAPPIST-1. 

\chapter{The Active Latitudes of HAT-P-11: Evidence for a Solar-like Dynamo} \label{chap:h11}

In the Introduction we introduced the solar dynamo and activity cycle, and how magnetic activity is expressed on the solar surface in the form of sunspots. The physical mechanisms which give rise the solar dynamo and activity cycle surely act on other Sun-like stars, as well, though our ability to study the latitude- and size distributions of starspots is hindered by their great distance from the Earth. Even with the worlds largest telescopes and interferometry, it is not yet possible to directly spatially resolve starspots in images of nearby stars. Thus, indirect methods must be employed.

In this chapter, we will use a transiting exoplanet to study the starspots of a star seven million times farther away from the Earth than the Sun, the K4 dwarf star HAT-P-11. The exoplanet HAT-P-11 b transits its host star from near the south stellar rotational pole to north pole, across many latitudes. As the planet blocks out starlight during transit events, it passes over dark starspots. When the planet occults a quiescent region of the stellar photosphere, up to 0.3\% of the starlight is obscured by the planet. When the planet occults a dark starspot, less light goes missing than during an occultation of a bright region -- imprinting a positive flux excursion onto the transit light curve. The amplitude of the change in flux during a spot occultation is proportional to the size of the spot and the contrast of the spot relative to the mean photosphere. The timing of the spot occultation encodes the latitudes of the occulted spots. Using this technique, we can reconstruct a map of the starspots of HAT-P-11 during transit events over the four-year \kepler mission. 

This chapter was first published in the Astrophysical Journal \citep{Morris2017a}.

\input{stsp_hat_p_11/accepted.tex}

\chapter{The Chromospheric Activity of HAT-P-11: an Unusually Active, Planet-hosting K Dwarf} \label{chap:h11sindex}

In the previous chapter, we opened a window into the stellar magnetic activity of the K4 dwarf HAT-P-11 using its transiting exoplanet. We found that HAT-P-11's starspots are distributed in latitude similar to sunspots, with most spots lying near $\pm15^\circ$ latitude. We found that the starspots have similar physical sizes to sunspots at solar maximum. Yet there are several questions that were left unanswered in the previous chapter. From the four years of \kepler photometry alone, it was not possible to build an analog to the solar butterfly diagram, showing the evolution of spot latitudes as a function of time; so one question that remains is -- how long is the full duration of the stellar magnetic activity cycle of HAT-P-11? 

Another unanswered question in the previous chapter is whether or not HAT-P-11 is ``normal''? In other words, for  planet-hosting stars of similar rotation periods and spectral types, is the solar-like activity of HAT-P-11 typical? We showed that the starspots of HAT-P-11 were similar sizes and distributed with latitude in a similar fashion to sunspots, and found that the typical spot coverage on HAT-P-11 was 100x greater than the Sun's. We now ask the question: is that activity level common among similar stars?

This chapter was first published in the Astrophysical Journal \citep{Morris2017b}.

\input{sindex/arxiv.tex}

\chapter{Stellar Properties of Active G and K Stars and Exploring the Connection Between Starspots and Chromospheric Activity} \label{chap:freckles}

\chapter{Spotting Stellar Activity Cycles with Gaia} \label{chap:gaia}

In the previous chapters, we have explored the affects of stellar activity on photometry and spectroscopy of active stars. There is a third dimension, in addition time and wavelength, that we can use to probe stellar activity -- astrometry. Stars are \textit{not} point sources, and as starspots rotate into and out of view, the astrometric centroid of a star will appear to shift slightly in response to the localized flux deficits. The astrometric precision required to measure this type of starspot-induced centroid shift is staggering, but may be within reach for the revolutionary ESA \gaia mission, as we will see in this chapter.

This chapter was first published in the Astronomical Journal \citep{Morris2018b}.

\input{gaia/arxiv.tex}

\chapter{Are Starspots and Plages Co-Located on Active G and K Stars?} \label{chap:nephelion}

In Chapter~\ref{chap:h11} we discovered that the latitude and radius distributions of starspots on HAT-P-11 are similar to the Sun's, and in Chapter~\ref{chap:h11sindex} we found that the chromospheric activity of HAT-P-11 is somewhat elevated compared to stars of similar rotation periods and spectral types. 

We focused on the dark starspots in Chapter~\ref{chap:h11} rather than the bright faculae which are probably present on the star for two reasons: (1) typical faculae have smaller flux contrast in the \kepler band compared to the mean photosphere than a typical starspot; and (2) our light curve inversion software \stsp assumes circular active regions with fixed contrast, which is a better approximation for starspots than it is for regions of faculae, which are often extended and increase in contrast towards the limb of the star. 

In this chapter, we attempt to measure the presence of faculae \textit{and} starspots with simultaneous observations of the \ion{Ca}{2} H \& K emission features -- which are brightest when faculae are facing the observer -- and photometric monitoring in red/optical bandpasses -- which is darkest when the most spotted face of the star faces the observer. 

This chapter was first published in the Astronomical Journal \citep{Morris2018g}. 

\input{nephelion/main.tex}

\chapter{Possible Bright Starspots of TRAPPIST-1} \label{chap:trappist1_bright}

In the previous chapter, we found that starspots and faculae are likely spatially associated with one another on the active young G star KIC 9652680, while we did not find a convincing correlation between the total flux and chromospheric emission for two mid-K stars in Praesepe, EPIC 211928486 and EPIC 211966629. Despite their broad range in rotation periods, in the broadest sense, these three stars are all ``Sun-like'' stars in that they have similar masses to the Sun, and have both convective and radiative zones, providing for the radiative-convective interface called the tachocline that lies at the heart of the $\alpha\Omega$ dynamo paradigm (refer back to Section~\ref{sec:solardynamo} for a brief review of the solar dynamo). 

In this chapter we seek to infer the stellar surface inhomogeneities of a likely very different star -- TRAPPIST-1 \citep{Gillon2016,Gillon2017}. TRAPPIST-1 is an M8 dwarf, and thus should be fully-convective. Lacking a radiative-convective boundary, the dynamo that powers magnetic activity may be significantly different from the one that operates within the Sun. 

This chapter was first published in the Astronomical Journal \citep{Morris2018c}.

\input{trappist1_bright/arxiv.tex}

\part{Exoplanet Characterization} \label{part:planets}
\begin{quote}
{\it Lo, the darting bowling orb!

Lo, the brother orbs around, all the clustering suns and planets,

 All the dazzling days, all the mystic nights with dreams,

Pioneers! O pioneers!}
\hfill --- Walt Whitman
\end{quote}

To date, the most productive exoplanet discovery and characterization method has been the transit technique. For systems where the planet appears to eclipse its host star, a temporary dip in brightness of the host star is detectable with precision photometry \citep{Charbonneau2000}. The dip in brightness is periodic and occurs once per orbit, allowing us to measure the planet's orbital period and semi-major axis via Kepler's law. The amplitude of the dip in brightness is proportional to the cross-sectional area of the planet to the star, $\Delta F/F = (R_p/R_\star)^2$, allowing us to measure the planet radius relative to the stellar radius.

In Chapter~\ref{chapter:apo}, we will discuss observations of transits of several planets gathered with the ARC 3.5 m Telescope at Apache Point Observatory, and what the transits tell us about: (1) the masses of exoplanets in Sections~\ref{sec:traptimes} and \ref{sec:koinet}; and (2) the spots of the host star in Section~\ref{sec:hatspot}.

\chapter{Ground-based Transit Photometry at Apache Point Observatory} \label{chapter:apo}

\section{Transit times of TRAPPIST-1 b and c with ARCTIC} \label{sec:traptimes}

This section was first published in Research Notes of the American Astronomical Society \citep{Morris2018a}.

\input{trappist1_rn/arxiv.tex}

\section{Holographic diffuser-assisted spot occultation photometry of HAT-P-11} \label{sec:hatspot}

This section was first published in Research Notes of the American Astronomical Society \citep{Morris2018d}.

\input{hat11rn/main.tex}

\section{KOINet Transit Timing Network} \label{sec:koinet}

The results of this section were first published in Astronomy \& Astrophysics \citep{vonEssen2018}.

\input{koinet/main.tex}

\part{The effects of stellar activity on planet characterization}

\begin{verse}
{\it Weep not, child,\\
  Weep not, my darling,\\
  With these kisses let me remove your tears,\\
  The ravening clouds shall not long be victorious,\\
  They shall not long possess the sky, they devour the stars only in apparition,\\
  Jupiter shall emerge, be patient, watch again another night, the Pleiades shall emerge,\\
  They are immortal, all those stars both silvery and golden shall shine out again,\\
  The great stars and the little ones shall shine out again, they endure,\\
  The vast immortal suns and the long-enduring pensive moons shall again shine.}
\hfill --- Walt Whitman
\end{verse}

In Part~\ref{part:activity} we discussed observations of stellar magnetic activity on a variety of stars, and in Part~\ref{part:planets} we discussed observations of transiting exoplanets and what they transmit about their masses and their host star's surfaces. These investigations were crucial for setting the stage for studying the effects of stellar magnetic activity on exoplanet characterization, a topic of ever-increasing importance in the age of the James Webb Space Telescope and thirty-meter-class telescopes, which will inevitably make advances in stellar astrophysics on their way to measuring properties of exoplanets.  

In this Part, we explore the effects of stellar activity on characterization of exoplanets. In Chapter~\ref{chapter:robin} we introduce the ``self-contamination'' transit light curve parameterization that can account for bright or dark starspots which contaminate transit observations, and apply the technique to observations of ten transiting exoplanets. In Chapter~\ref{chapter:robin_trap} we apply the self-contamination technique to recent \spitzer photometry of the TRAPPIST-1 planets to look for evidence of starspot contamination. Finally in Chapter~\ref{chapter:jwst_nirspc}, we bring together our expertise on stellar activity and transiting exoplanets in order to generate simulated observations of transiting exoplanets which orbit active stars for observations with JWST/NIRSpec.  

\chapter{Robust Transiting Exoplanet Radii in the Presence of Starspots from Ingress and Egress Durations} \label{chapter:robin}

In Chapter~\ref{chap:h11} we used the transiting exoplanet HAT-P-11 to characterize the starspots that it occulted. In that chapter, we had to take extra care to measure the photometric transit depth from the \kepler photometry, because spot occultations occurred frequently, contaminating most transits towards shallower depths. In this chapter, we address this problem head-on: suppose you had a star with a dense band of starspots or a bright ring beneath the transit chord of the exoplanet? Such a surface inhomogeneity might be incorrectly interpreted as a smaller or larger planet; but we will find that the ingress and egress durations of transiting exoplanets provide an independent check on the radius of the exoplanet, which is not contaminated by starspots in the transit chord.    

This chapter was first published in the Astronomical Journal \citep{Morris2018e}.

\input{robin/main.tex}

\chapter{Robust Transiting Exoplanet Radii: Revisiting TRAPPIST-1} \label{chapter:robin_trap}

In the previous chapter we developed and applied the ``self-contamination'' technique used to analyze transit light curves of planets which occult spotted stars. The self-contamination simply modifies the \citet{Mandel2002} transit light curve equation such that the transit depth is disconnected from the ratio of planet to stellar radii, allowing the depths of transit light curves to be contaminated by the presence of starspots, bright or dark, occulted or unocculted. 

In this chapter, we apply the self-contamination technique to a star we have already visited with great interest in this thesis: TRAPPIST-1. As we explored in Chapter~\ref{chap:trappist1_bright}, bright spots may be required to explain the \kepler and \spitzer photometry of TRAPPIST-1. Knowing the true properties of the stellar activity of TRAPPIST-1 is key to understanding the transmission spectra of its exoplanets. Thus in this chapter, we use recent \spitzer transit observations of TRAPPIST-1 b-h to probe the stellar surface inhomogeneities of TRAPPIST-1 with the self-contamination technique.

This chapter was first published in the Astrophysical Journal Letters \citep{Morris2018f}.

\input{trappist1_robin/main.tex}

\chapter{The Limits of Transit Characterization with JWST/NIRSpec due to Stellar Variability: The Cases of TRAPPIST-1, Kepler-296 and Kepler-62.} \label{chapter:jwst_nirspc}

\chapter{Conclusions}

\begin{verse}
{\it When I heard the learn'd astronomer,\\
When the proofs, the figures, were ranged in columns before me,\\
When I was shown the charts and diagrams, to add, divide, and measure them,\\
When I sitting heard the astronomer where he lectured with much applause in the lecture-room,\\
How soon unaccountable I became tired and sick,\\
Till rising and gliding out I wander'd off by myself,\\
In the mystical moist night-air, and from time to time,\\
Look'd up in perfect silence at the stars.\it}
\hfill --- Walt Whitman
\end{verse}

\section{Outlook for exoplanet detection with TESS and ground-based follow-up}

The TESS mission will survey 200,000 of the brightest stars in the sky for evidence of transiting exoplanets \citep{Ricker2014}. Most planet candidates will only be observed by TESS for 27 consecutive days, so only a handful of transits will be recorded in the primary mission. It will be the task of ground-based follow-up observers to monitor planet candidates produced by TESS to confirm the planetary nature of these candidates by planetary multiplicity, transit timing variations, or radial velocity campaigns. 

In Chapter~\ref{chapter:apo} (Section~\ref{sec:hatspot}), we showed that a holographic diffuser enables high precision photometry from the ground \citep[see also][]{Stefansson2017,Stefansson2018} -- diffuser photometry may be a key asset for follow-up of TESS targets. The diffuser transforms the point-spread function of a star from a nearly Gaussian distribution with a tight radial spread of only a few pixels to a broad top-hat shaped distribution covering an area of thousands of pixels. For bright stars, spreading the light out over many pixels is advantageous for several reasons: (1) it enables longer exposures, which can become long enough to average over atmospheric scintillation noise; (2) it makes the photometry less sensitive to fine pointing errors and intra-pixel sensitivity variations, which are significant when most of the stellar flux is landing on only a handful of pixels; and (3) the photometry is also less sensitive to inter-pixel sensitivity variations (or flat-field precision), since the flux is spread roughly evenly over many pixels. This last point is one of the key advantages in favor of diffuser photometry over defocused photometry, where the flux is often concentrated unevenly onto a few pixels. 

\subsection{Stellar activity of TESS host stars}

Diffuser photometry is not only useful for measuring transits and and transit timing variations, but the uniquely high precision diffuser photometry is well-suited to measuring spot occultations from the ground \citep{Morris2018d}. Should more stars be discovered like HAT-P-11 which show frequent spot occultations, ground-based diffuser photometry may allow us to measure properties stellar magnetically active regions over very long baselines, for example, throughout stellar activity cycles. Also exciting is the possibility of expanding upon the monochromatic photometry of TESS by observing spot occultations from the ground in a different bandpass. For example, we can collect ground-based photometry of HAT-P-11 in the SDSS $u$ band while TESS observes in its redder bandpass, allowing us to use the simultaneous observations to roughly estimate the color, and therefore the temperature, of starspots on HAT-P-11 (or other stars). 

\section{Outlook for exoplanet characterization with JWST}

The James Webb Space Telescope (JWST) may provide some of the clearest views into exoplanet atmospheres and their masses to date, using transmission spectroscopy and transit timing variations (TTVs). 


\printendnotes

%
% ==========   Bibliography
%
% \nocite{*}   % include everything in the uwthesis.bib file
\bibliographystyle{apj}
\bibliography{uwthesis,./intro/bibliography,./stsp_hat_p_11/bibliography,./sindex/sindex_bibliography,./gaia/gaia_bib,./trappist1_bright/t1bright_bib,./trappist1_rn/t1rn_bib,./hat11rn/bibliography,./astroplan/bibliography,./koinet/bibliography,./robin/bibliography,./trappist1_robin/bibliography,./nephelion/bibliography,./howard/bibliography}

%
% ==========   Appendices
%
\appendix
\raggedbottom\sloppy

\pagebreak
In the previous Parts, thousands of lines of code were written to facilitate each scientific effort, all of which are available on GitHub\footnote{\url{http://github.com/bmorris3}}. In the appendices, we highlight two open source, open development coding efforts which may be of use to the broader community. In Appendix A we introduce {\tt astroplan}, an astropy-affiliated package for observation planning and scheduling in Python. In Appendix B we introduce {\tt aesop}, a Python package for analysis of echelle spectra collected with the ARC Echelle Spectrograph (ARCES) at Apache Point Observatory (APO). 

\chapter{astroplan: an Open Source Observation Planning Package in Python} \label{chap:astroplan}

This chapter was first published in the Astronomical Journal \citep{astroplan}.

\input{astroplan/arxiv.tex}

\chapter{aesop: The ARC Echelle Spectrograph Observation Pipeline} \label{chap:aesop}

This chapter was first published in the Journal of Open Source Software in collaboration with Trevor Dorn-Wallenstein \citep{aesop}.

The ARC Echelle Spectroscopic Observation Pipeline, or {\tt aesop}, is a high resolution spectroscopy software toolkit tailored for observations from the Astrophysics Research Consortium (ARC) Echelle Spectrograph mounted on the ARC 3.5 m Telescope at Apache Point Observatory. {\tt aesop} picks up where the traditional IRAF reduction scripts leave off, offering an open development, object-oriented Pythonic analysis framework for echelle spectra.

Basic functionality of {\tt aesop} includes: (1) blaze function normalization by polynomial fits to observations of early-type stars, (2) an additional/alternative robust least-squares normalization method, (3) radial velocity measurements (or offset removals) via cross-correlation with model spectra, including barycentric radial velocity calculations, (4) concatenation of multiple echelle orders into a simple 1D spectrum, and (5) approximate flux calibration.

Some handy additional utilities include methods for download PHOENIX model spectra \citep{Husser2013}, and methods for measuring the Mount Wilson Observatory-calibrated CaII H & K ``S'' index \citep{Morris2017b}. aesop was built from the Astropy package-template, and thus includes self-building documentation and continuous integration \citep{Astropy2018}.
 
 
% ========== Appendix A

\vita{Brett Morris grew up on Long Island, in New York. He graduated from the University of Maryland with BS in Astronomy with High Honors, and a BS in Physics. He worked as a research associate at NASA's Goddard Space Flight Center before attending the University of Washington for his dual-title PhD in Astronomy and Astrobiology.}

\end{document}
