
\begin{verse}
{\it When I heard the learn'd astronomer,\\
When the proofs, the figures, were ranged in columns before me,\\
When I was shown the charts and diagrams, to add, divide, and measure them,\\
When I sitting heard the astronomer where he lectured with much applause in the lecture-room,\\
How soon unaccountable I became tired and sick,\\
Till rising and gliding out I wander'd off by myself,\\
In the mystical moist night-air, and from time to time,\\
Look'd up in perfect silence at the stars.\it}
\hfill --- Walt Whitman
\end{verse}

\chapter{Conclusions}

\section{Summary}

In Part~\ref{part:solaractivity} we studied the magnetic activity of the Sun, and its impact on observations of Earth-like exoplanets orbiting Sun-like stars. In Chapter~\ref{chap:solarbenchmark} we reconstructed the rotational modulation of the Sun over 68 years of observations using archival sunspot records. In Chapter~\ref{chap:plato} we found that stellar oscillations and granulation are a significant source of astrophysical noise in measuring exoplanet radii with precision, with applications to observations by the PLATO mission. In Chapter~\ref{chap:astero} we exploited a numerical technique for efficiently generating long-term light curves of Sun-like oscillators including the effects of $p$-mode oscillations, granulation and supergranulation. In conclusion, Part~\ref{part:solaractivity} built a foundational understanding of solar magnetic activity and variability, setting the stage for studies of distant stellar surfaces.

In Part~\ref{part:activity} we presented detailed studies of the magnetic activity of HAT-P-11, a K4V with a transiting hot-Neptune in a fortuitous orbit. The planet transits from near the south pole of the star to the north pole, which allowed us to resolve starspots as a function of latitude on the stellar surface in Chapter~\ref{chap:h11}. The spots reveal that the star likely has a Sun-like dynamo, producing spots with a similar latitudinal distribution, though the spot coverage is two orders of magnitude larger. In Chapter~\ref{chap:h11sindex} we studied the long-term chromospheric activity of HAT-P-11 and found that the star likely has a $\gtrsim11$ year activity cycle similar to the Sun's, and we provide evidence that the close-in planet could induce the enhanced stellar activity that we observed. 

In Chapter~\ref{chap:nephelion} we investigated whether plages regions are spatially associated with dark starspots on a young G star and two K stars, and find evidence for a Sun-like association between starspots and plages for the G star, but no such evidence for the K stars, which seem to have global chromospheric activity not closely associated with the dark starspots driving rotational modulation. We measured the stellar temperatures, surface gravities, and spot covering fractions for a sample of bright stars in Chapter~\ref{chap:freckles}, and find that chromospheric activity weakly correlates with spot covering fraction, and that high precision color priors are crucial to measuring sensible spot coverages. 

We turned our attention to the magnetic activity of the very lowest mass planet-hosting star, TRAPPIST-1, in Chapter~\ref{chap:trappist1_bright}. Using \kepler and \spitzer photometry, we reasoned that TRAPPIST-1 likely has bright regions driving the stellar variability observed at optical wavelengths but not observed in the near infrared. We also found evidence that the presence of the bright regions seems to be associated with flare occurrence, perhaps indicating that the apparent rotational modulation observed at optical wavelengths may not be rotational modulation after all. Finally, we looked to astrometry for a novel measurement of extreme stellar magnetic activity by its influence on stellar astrometric centroids in Chapter~\ref{chap:gaia}.  

In Part~\ref{part:synthesis} we brought together our knowledge of stellar magnetic activity and exoplanet characterization to prepare for observations of planets orbiting active host stars. In Chapter~\ref{chapter:robin} we considered a ``worst-case scenario'' where starspots affect our measurements of exoplanet radii, and found that robust exoplanet radii can still be recovered measuring the durations of ingress and egress. In Chapter~\ref{chapter:robin_trap} we applied this technique to new \spitzer observations of TRAPPIST-1 and found no evidence for significant contamination of the transit chords of the TRAPPIST-1 planets by stellar activity. Lastly, in Chapter~\ref{chapter:jwst_nirspc} we prepared for observations of the systems of potentially habitable planets of TRAPPIST-1, Kepler-62 and Kepler-296 with the James Webb Space Telescope (JWST) using NIRSpec's in Prism mode. We determined that TRAPPIST-1 may be one of the best known systems for follow-up with James Webb due largely to its brightness and the small size of its host star. We outlined a plan for measuring the mass of the potentially habitable planet Kepler-62 f with precision as good as 10\%, and showed that transmission spectroscopy of the TRAPPIST-1 planets can discriminate between different climates in the absence of clouds. 

\section{Concluding Thoughts}

In this dissertation we have used a variety of observational techniques to constrain the heterogeneity of stellar surfaces due to magnetic activity. Looking towards the coming decade, efforts to precisely characterize stellar surface inhomogeneities will become increasingly important as we invest significant resources in collecting spectra of small exoplanet atmospheres. \citet{Cowan2015}, for example, suggested that perhaps the most efficient way to allocate JWST observation time towards the study of exoplanet atmospheres is to perform a ``narrow and deep'' survey, limiting our observations to only a handful of exoplanets studied over a long period of time. A similar strategy may be carried out by extremely large telescopes (ELTs) on the ground. Accurate interpretation of these expensive spectra must be built upon a comprehensive understanding of the stellar surfaces which illuminate each exoplanet, as a function of wavelength and time. We have summarized these recommendations in the 2020 decadal white paper by \citet{Rackham2019}. 

\subsection{The Sun as a Star}

We are quite certain that the efforts in Part~\ref{part:solaractivity} to characterize the Sun as a star, and to predict the effects of Sun-like magnetic activity and variability on exoplanet characterization, are only parts of a prologue to a much longer story written for the exoplanet community about the only easily-resolved star from Earth. The Sun will always be the best-studied star in the sky, and for the foreseeable future, it will remain the only known host star to an inhabited exoplanet. As such, the Sun and the heliophysics it enables should always hold a special place in the minds of exoplaneteers as we study distant stars. 

\subsection{Magnetic activity of Sun-like Stars}

Often in studies of exoplanets and habitability, many stars are lumped together under the term ``Sun-like stars'', generally encompassing spectral types F through K. In this dissertation, we have presented evidence that G and K main-sequence stars may in fact exhibit similar magnetic activity to the Sun's. In Chapter~\ref{chap:h11} we showed that the latitude and radius distributions of starspots on the K4 dwarf HAT-P-11 are consistent with the Sun's at solar maximum, though the spot coverage is two orders of magnitude greater than solar. In Chapter~\ref{chap:h11sindex} we showed that the activity cycle duration of HAT-P-11 may be similar to the Sun's. In Chapter~\ref{chap:nephelion} we showed that the bright faculae and starspots of a young G star may be physically associated with dark starspots, as they are on the Sun. From the perspective of stellar magnetic activity, it appears that the catch-all term ``Sun-like stars'' is justified for both G and K type main-sequence stars, with the important caveat that the quantity of starspots likely increases with later spectral type. This may be an important justification for observations in the coming decades as more planets are discovered orbiting G and K stars by TESS and PLATO, and since some have advocated that K dwarfs may be the ideal candidate host stars for searches for biosignatures in exoplanet atmospheres \citep[see, for example,][]{Arney2019}.

\subsection{HAT-P-11 as a Benchmark for Active Stars}

Rarely have starspot analyses focused as much effort on characterizing the spots of a single star as we have spent on HAT-P-11 (with one glaring exception being the Sun). Most of the tests developed in this dissertation have been applied to that fortuitous system, which is perhaps triply lucky given its relative brightness ($V=9.5$), its highly-inclined exoplanet orbit, and its ample spot coverage. It is our sincerest hope that HAT-P-11 can act as a benchmark star, against which other active G and K stars can be compared. Should we be lucky enough to discover a habitable planet orbiting a nearby mid-K star, we will be able to analyze spectroscopy of that planet's atmosphere with informative prior expectations for the stellar activity of its host thanks to HAT-P-11. 

\subsection{Photometric variability of TRAPPIST-1}

The magnetic activity of TRAPPIST-1 will continue to beguile astronomers after this dissertation has concluded. We presented one of the first multi-color photometric analyses of the stellar variability of TRAPPIST-1 in Chapter~\ref{chap:trappist1_bright}, with an additional (inconclusive) supporting analysis in Chapter~\ref{chapter:robin_trap}, which illuminated several interesting properties of this important exoplanet host star. The variability observed at optical wavelengths and the lack of variability in the near-infrared can be described by a few bright, hot regions on the stellar surface. We note an intriguing correlation between the emergence of bright regions and the occurrence of flare events, which perhaps suggests that the variability of the star is due to the emergence and decay of active regions temporally associated with the flares, rather than rotational modulation. This may also explain the apparently discrepant rotation periods of the star in the literature \citep{Gillon2016,Luger2017}. 

\subsection{Prospects for future applications of the methods presented}

We have developed and refined a number of techniques in this dissertation which allow observers to make the most of the information contained in transit light curves of exoplanets, particularly with respect to characterizing the stellar magnetic activity of exoplanet host stars. One useful way of discussing the utility of the techniques presented here, popularized by Michael Gully-Santiago, is in terms of the ``scalability'' of the method, i.e.~how widely we can hope to apply each method, and the ``fidelity'' of the method, i.e.~how informative and quantitative are the results produced by each method. For example, the spot occultation modeling technique applied to HAT-P-11 in Chapter~\ref{chap:h11} allowed us to resolve small starspots and to identify their positions on the stellar surface, so it may be said that these observations were ``medium-fidelity''. However the MCMC inference technique was computationally expensive, and the ability to map the latitude distribution of starspots is applicable only to transiting exoplanets with highly-inclined orbits, so the spot occultation technique has ``low-scalability.'' We are optimistic that analogs of HAT-P-11 and Kepler-17 will be discovered in TESS and PLATO observations, which will enable us to apply this method to a handful of stars. While the number of spot-occultation targets will be small, the unique ability of spot occultations to test predictions of stellar magnetic dynamo theory \citep[see, for example,][]{Gilman2018} promises that future efforts to expand upon Chapter~\ref{chap:h11} will be worthwhile. 

Our experiment with the TiO absorption technique presented in Chapter~\ref{chap:freckles} showed that precise characterization of starspots requires telescopes with larger apertures and/or spectrographs with higher resolutions than the $R=31,500$ ARC Echelle Spectrograph on the ARC 3.5 m Telescope at APO, even for bright stars; thus we deem the TiO absorption method medium-fidelity and low-scalability. However, we note that it is already common practice to collect high resolution, high signal-to-noise spectra of exoplanet host stars, and thus it may be possible to mine the wealth of spectra of exoplanet host stars already obtained for starspot analyses.  

The two-color photometric analysis of TRAPPIST-1 in Chapter~\ref{chap:trappist1_bright}, along with searches for correlations between stellar flux and flare occurrence, must be repeated for more M8V stars to understand TRAPPIST-1's purported magnetic activity. Some spectroscopic evidence is accruing that the bright spots we hypothesized in Chapter~\ref{chap:trappist1_bright} may be real \citep[see, for example,][]{Ducrot2018, Wakeford2019}, but the conflicting reports for TRAPPIST-1's rotation period remain unresolved in the literature. Since TESS is likely to discover the best exoplanets amenable to characterization with JWST, and TESS observes in the optical while JWST observes in the infrared, we will likely have access to photometry in two colors for the most promising exoplanet hosts. Thus we note that the spot modulation analysis of Chapter~\ref{chap:trappist1_bright} has high-scalability and low-fidelity. 

The self-contamination technique in Chapters~\ref{chapter:robin} and \ref{chapter:robin_trap} is applicable to all transiting exoplanet observations, though it is most informative when applied to short-cadence photometry, such as photometry from the PLATO mission (with 45-second cadence). However, the inferences it makes about the properties of starspots are qualitative, ranking the technique: high-scalability and low-fidelity. 

This brief survey of the starspot analysis methods presented in this dissertation demonstrates that each individual technique is incapable of answering our many questions about the starspots of a given star. Perhaps our greatest hope is to combine all of the above techniques to construct the most detailed picture of the stellar magnetic activity of a star. In addition to our multi-pronged approach to uncovering the starspots of HAT-P-11, another comprehensive analysis of note was carried out for the star LkCa 4 by \citet{Gully2017}. We hope to see the establishment of more benchmark active stars like HAT-P-11 and LkCa 4 discovered and characterized in the near future.


\section{Future Work}

Moving forward, there are several questions that will define the scope of my future research, including: 
\begin{enumerate}
\item Do TESS two-minute cadence observations provide any constraints on the radii of the smallest exoplanets through the durations of ingress and egress?
\item Are there HAT-P-11-like systems with large amplitude spot occultations in TESS two-minute cadence observations? Can we make inferences about stellar dynamos as a function of stellar mass using spot occultations? Can we follow-up these targets using ground-based observations?
\item What newly discovered small exoplanets may take the lead as the most promising candidates for atmospheric and mass characterization with JWST? 
\item What existing and future spectroscopic observations can be taken to constrain the spot coverages of planet-hosting stars? 
\item Can we make use of future data releases of ESA Gaia astrometry to detect the photometric centroid offsets of stars due to their magnetic activity? 
\item Is the photometric variability of TRAPPIST-1 driven by rotational modulation, or something else? 
\end{enumerate}

In the sections that follow, we expand upon each of these questions and provide strategies for answering them.

\subsection{Exoplanet detection and characterization with TESS and ground-based follow-up}

The Transiting Exoplanet Survey Satellite (TESS) will observe 200,000 stars at 2 minute cadence, providing a \kepler-like photometric database for the brightest stars distributed throughout the entire sky. Unlike \kepler, most TESS stars will be observed for only 27 consecutive days, compared to the four years of \kepler photometry. As a result of these short observing windows, when it comes to characterizing exoplanet host star magnetic activity, we will have to make more with less space-based photometry, and lean more heavily on ground-based follow-up observations. For TESS targets to be amenable to follow-up characterization with JWST, they must have either little or well-characterized stellar activity. I am prepared to provide a multi-pronged approach to characterizing the activity of exoplanet host stars observed with TESS, in preparation for characterization by JWST. I will achieve this goal with TESS photometry, and ground-based follow-up observations.

The deepest insights into stellar magnetic activity of exoplanet host stars come from the transit light curves of exoplanets. There are two primary techniques that I plan to use to measure the activity of exoplanet host stars using the publicly available TESS photometry: the ``self-contamination'' technique, and spot occultation mapping. 

\subsubsection{Robust radii for small exoplanets via the self-contamination technique}

We usually measure the radii of transiting exoplanets from the transit depth, which are given by the ratio of cross-sectional areas of the planet and star. However, if a star has dark starspots distributed throughout the transit chord, the transit depth will be biased toward smaller values, and thus the inferred planet radius will appear smaller if these spots are unaccounted for. In \citet{Morris2018f, Morris2018g} we reparameterize the transit light curve to account for self-contamination by photospheric inhomogeneities by splitting the parameter $R_p/R_\star$ into two parameters: one for the radius ratio, which controls the duration of ingress and egress, and another which measures the possibly contaminated transit depth. The radius recovered from measurement of the ingress/egress duration can recover the true radii of planets transiting spotted stars with axisymmetric spot distributions, especially when asteroseismic observations are available. This will be critical for measuring robust radii of small exoplanets orbiting the brightest stars, which are the focus of the TESS mission, using the publicly available TESS photometry alone.

\subsubsection{Spot occultation mapping for precision dynamo studies}

The clearest window into small-scale magnetic activity of exoplanet host stars to date is spot occultation mapping, which reveals the positions and sizes of starspots as exoplanets occult them during transits \citep{Sanchis-Ojeda2011}. When coupled with Rossiter-McLaughlin observations which constrain the spin-orbit alignment of an exoplanet system, spot occultations can reveal the latitudes and longitudes of spots on the stellar surface, allowing for direct comparison with solar activity \citep[][see Chapter~\ref{chap:h11}]{Winn2010, Morris2017a}. I will search the TESS photometry of exoplanet host stars for every spot occultation observed during by TESS mission. For a subset of the stars with Rossiter-McLaughlin observations, I will reconstruct the latitude/longitude distributions of the spots, enabling precision comparisons between stellar and solar activity, as I did in Chapter~\ref{chap:h11}. 

Holographic diffusers enable high precision photometry from the ground \citep{Stefansson2017,Stefansson2018,Morris2018a} -- diffuser photometry may be a key asset for follow-up of TESS targets. The diffuser transforms the point-spread function of a star from a nearly Gaussian distribution with a tight radial spread of only a few pixels to a broad top-hat shaped distribution covering an area of thousands of pixels. For bright stars, spreading the light out over many pixels is advantageous for several reasons: (1) it enables longer exposures, which can become long enough to average over atmospheric scintillation noise; (2) it makes the photometry less sensitive to fine pointing errors and intra-pixel sensitivity variations, which are significant when most of the stellar flux is landing on only a handful of pixels; and (3) the photometry is also less sensitive to inter-pixel sensitivity variations (or flat-field precision), since the flux is spread roughly evenly over many pixels. This last point is one of the key advantages in favor of diffuser photometry over defocused photometry, where the flux is often concentrated unevenly onto a few pixels.

\subsubsection{Stellar activity of TESS host stars}

Since TESS will observe many more late K and M type stars than \kepler \citep{Muirhead2018}, the mission will enable us to construct spot occultation maps for more stars across the low-mass end of the main sequence, allowing us to probe the behavior of stellar dynamos as a function of stellar mass. At present, small-scale spot maps are only available for a handful of stars, such as Kepler-17, Kepler-63, HAT-P-11 and the Sun \citep{Solanki2003,davenportthesis,sanchis-ojeda2013,Morris2017a}. TESS may open a window into the dynamos of low-mass stars, allowing us to determine how magnetic activity changes with rotation period, tachocline depth, Rossby number, or other fundamental stellar parameters which are thought to influence stellar dynamos \citep{Berdyugina2005, Gilman2018}. 

Diffuser photometry is not only useful for measuring transits and and transit timing variations, but the uniquely high precision diffuser photometry is well-suited to measuring spot occultations from the ground \citep{Morris2018d}. Should more stars be discovered like HAT-P-11 which show frequent spot occultations, ground-based diffuser photometry may allow us to measure properties stellar magnetically active regions over very long baselines, for example, throughout stellar activity cycles. Also exciting is the possibility of expanding upon the monochromatic photometry of TESS by observing spot occultations from the ground in a different bandpass. For example, we can collect ground-based photometry of HAT-P-11 in the SDSS $u$ band while TESS observes in its redder bandpass, allowing us to use the simultaneous observations to roughly estimate the color, and therefore the temperature, of starspots on HAT-P-11 (or other stars). 

\subsection{Exoplanet characterization with JWST}

The TESS planet yield, in addition to ground-based surveys like SPECULOOS \citep{Delrez2018b} and SAINT-Ex, should produce a handful of planets which are ideal for characterization with the James Webb Space Telescope (JWST). JWST may provide some of the clearest views into exoplanet atmospheres and their masses to date using transmission spectroscopy and transit timing variations (TTVs), as we discussed in Chapter~\ref{chapter:jwst_nirspc}. 

While stellar magnetic activity, granulation and oscillations will enforce fundamental noise floors on the precision of JWST observations of exoplanets, we are optimistic that some targets may be bright and quiet enough to measure the masses of potentially habitable (cool) planets via TTVs, and to measure the presence or lack of an atmosphere for some close-in (hot) rocky planets. 

If we are lucky enough to have a system where both techniques can be applied to the same planet, in principle, we can measure the bulk density of the planet, {\it and} the atomic and molecular constituents of its atmosphere. For the first time, we may be able to say something about the bulk chemistry of an exoplanet's interior as well as basic properties of its atmosphere like the presence or lack of clouds. In the case of a cloud-free atmosphere, we might also be able to discern between different climates on the planet via its atmospheric chemistry. 

\subsection{Ground-based Spectroscopic Follow-up of Exoplanet Hosts}

In Chapter~\ref{chap:freckles}, we showed that the resolution and signal-to-noise of ARCES spectra from the ARC 3.5 m Telescope at Apache Point Observatory are insufficient to detect starspot coverages $\lesssim10\%$. We can improve upon this effort in the future, for example, by turning to archival observations from W.M.~Keck Observatory. There is a valuable database of high resolution, high signal-to-noise spectra for 1305 Kepler Objects of Interest collected by the California-Kepler Survey \citep[CKS:][]{Petigura2017,Johnson2017}. This treasure-trove of spectra may contain evidence for starspot coverage due to absorption by TiO on exoplanet host stars, for example, at 7054 and 7125 \AA\ which are covered by the CKS archival spectra. 

\subsection{Gaia Observations of Stellar Magnetic Activity}

In Chapter~\ref{chap:gaia} we described a technique for observing stellar activity through photometric centroid offsets using the exquisite photometric precision afforded by ESA's Gaia mission. In the near future, Gaia data releases will provide constraints on the ``excess astrometric noise'' detected for each bright, nearby target. This excess noise will be used to probe stellar multiplicity and the presence of exoplanets, but it can also be used to search for extreme activity for some of the nearest stars. I will lead efforts to measure excess astrometric noise due to stellar magnetic activity on the nearby, bright stars that we suggested as high priority targets in Chapter~\ref{chap:gaia}. 

\subsection{The Activity of TRAPPIST-1}

Since the publication of Chapter~\ref{chap:trappist1_bright}, observations have led other groups to also conclude that there are likely bright, hot regions on the surface of TRAPPIST-1 \citep{Ducrot2018, Wakeford2019}. With upcoming observations of late-M stars like those of SPECULOOS and SAINT-Ex, it may be possible to determine whether or not these apparently bright regions are common, or if they are particular to TRAPPIST-1. It may also be possible to mine these observational databases of other stars for evidence of the flux-flare correlation described in Chapter~\ref{chap:trappist1_bright}. We are optimistic that by studying populations of late-M stars, it may be possible to identify which aspects of TRAPPIST-1's activity are due to bright spots, which are due dark spots, and whether or not the apparent rotational modulation of TRAPPIST-1 is in fact due to rotation rather than the evolution of active regions. We also note that observations with very high resolution spectrographs, like PEPSI at $R=270,000$, may make it possible to measure the stellar $v \sin i$ and thus the rotation period, even on a very late M dwarf \citep{Strassmeier2015}. With these observations we might hope to overcome the degenerate interpretations of the optical light curve of TRAPPIST-1, which may be the result of rotational modulation, or the evolution of active regions. 

\subsection{Outlook for astrobiology}

The aforementioned observations are the primary ingredients that go into determining the true habitability of an exoplanet. It is possible that JWST will usher us into a new era in the study of exoplanet habitability, leaving behind the rough definition of habitability based on the instellation, by using the moist and maximum greenhouse effects as boundaries of the habitable zone, and moving onward to identifying whether or not the observed surface and climate properties of a specific planet are consistent with a habitable planet or not. With careful consideration given to the properties of the host stars, we are poised to ask one of the paramount questions of astrobiology: where are the conditions for life met in the local Universe?
